### 2018-03-20 proposal/proposal.tex
Added proposal base tex file

\documentclass[12pt,a4paper]{article}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{float}
\usepackage{subcaption}
\usepackage{cleveref}
\usepackage{amsmath}
\usepackage{enumitem}

\newcommand{\scomment}[1]{{\bf{\color{red}{{[Suyash: #1]}}}}}
\newcommand{\acomment}[1]{{\bf{\color{blue}{{[Aman: #1]}}}}}
\newcommand{\fullref}[1]{\hyperref[{#1}]{\autoref*{#1} - \nameref*{#1}}} % One single link
\newcommand{\figref}[1]{\hyperref[{#1}]{\autoref*{#1}}} % One single link

\hypersetup{
colorlinks=true,
allcolors=blue,
}

\usepackage{geometry}
\geometry{
a4paper,
left=20mm,
top=30mm,
}

\title{COL772: Project Proposal}
\author{Suyash Agrawal (2015CS10262) \\ Aman Agrawal (2015CS10210)}

\begin{document}
\maketitle
\section{Text Classification}
    \bibliographystyle{unsrt}
    \bibliography{proposal}
\end{document}