### 2018-03-20 proposal/proposal.tex
added data section in project proposal

\documentclass[10pt,a4paper]{article}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{float}
\usepackage{subcaption}
\usepackage{cleveref}
\usepackage{amsmath}
\usepackage{enumitem}

\newcommand{\scomment}[1]{{\bf{\color{red}{{[Suyash: #1]}}}}}
\newcommand{\acomment}[1]{{\bf{\color{blue}{{[Aman: #1]}}}}}
\newcommand{\fullref}[1]{\hyperref[{#1}]{\autoref*{#1} - \nameref*{#1}}} % One single link
\newcommand{\figref}[1]{\hyperref[{#1}]{\autoref*{#1}}} % One single link

\hypersetup{
colorlinks=true,
allcolors=blue,
}

\usepackage{geometry}
\geometry{
a4paper,
left=19mm,
right=19mm,
top=20mm,
}

\title{COL772: Project Proposal}
\author{Suyash Agrawal (2015CS10262) \\ Aman Agrawal (2015CS10210)}

\begin{document}
\maketitle
\section{Goal}
    Our aim is to create a model that can learn to generate reply for tweets. Further, we would also experiment
    with other domains like email and dialogues.

\section{Dataset}
We plan to use the Marsan Ma’s curated twitter corpus of $800,000$ context-responses\cite{data:twitter-marson} for our baseline model testing. Later, we can actually use the twitter streaming API to mine tweets and their corresponding replies to create our own dataset in a day or so.\\
We can later extend the dataset by merging it with the Enron Email Dataset, which contains about $500k$ emails of $150$ employees of Enron Corporation\cite{data:enron}. \\
If, time permits, we can concatenate the Cornell Movie-Dialogs corpus \cite{data:cornell-movie} in addition to the Twitter corpus to train a better model.

\section{Literature}
    Recently, google released a paper ``Smart Reply''\cite{paper:smartReply}, which is their model for 
    email auto responder. Apart from this, there are various papers on text summarization\cite{paper:pointer}
    where we plan to take inspiration from.\\
    We have not found a paper that exactly does twitter replies and thus we would be writing code from scratch.
\section{Evaluation}
    For quantitative evaluation, we are currently planning on using the perplexity as our loss measure. The gold
    labels, would be a held out test data from the whole dataset. We would also be reporting the BLEU, METEOR and
    ROGUE metric of our predictions.\\
    For qualitative evaluation, we can just look at some randomly sampled results and judge their quality.

    \bibliographystyle{unsrt}
    \bibliography{proposal}
\end{document}